### 2018-08-09 CS761-RandomizedAlgorithms/assignments/final.tex
q6

%Problem 3
\newpage
\section{}
\paragraph{}
We throw $n^{0.99}$ balls into $n$ bins uniformly at random. Using union bound, the probability that a bin has at least $k$ balls, for $k$ chosen later, is at most 
$${n \choose k}(\frac{1}{n})^k.$$
Using the bound ${n^{0.99}\choose k} <(\frac{n^{0.99}e}{k})^k$ this probability is at most
$$(\frac{n^{0.99}e}{k})^k(\frac{1}{n})^k= \frac{e^k}{n^{0.0k}k^k}\leq \frac{e^k}{k^k}$$
where the last inequality follows since we will choose $k \geq 0$ and since $n\geq 1$ this implies $n^\frac{k}{100} \geq 1$.
Again invoking union bound
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq n\frac{e^k}{k^k} = n(\frac{e}{k})^k.$$ 
Since we want
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq \frac{1}{2}$$
we need to find $k$ such that
$$n(\frac{e}{k})^k \leq \frac{1}{2}.$$
%Problem 4
\newpage
\section{}
\paragraph{}
We always use addition, $+$, over the appropriate field (ie. $\F_n$ for inputs and $\F_m$ for outputs). Our algorithm is very simple. Given input $z$ choose $x \in \{0,\dots, n-1\}$ uniformly at random. Let $y = z-x$. Lookup $F(x)$ and $F(y)$ and return $F(x) + F(y)$.
\paragraph{}
If both $F(x)$ and $F(y)$ are not modified by our adversary then our return value satisfies
$$F(x) + F(y) = F(x+y) = F(z)$$
using the linearity of $F$ and the fact that $x+y = z$. So a necessary condition for our algorithm to fail is that one of $F(x)$ or $F(y)$ was modified by our adversary. For a uniform random element of the table, the probability of modification is at most $\frac{1}{5}$. Hence
$$Pr(F(x)\text{ or } F(y)\text{ was modified}) \leq Pr(F(x) \text{ was modified}) + Pr(F(y)\text{ was modified}) \leq \frac{2}{5}.$$
Therefore the probability of failure is at most
$$\frac{2}{5} < \frac{1}{2}.$$
Hence our algorithm succeeds with probability at least $\frac{1}{2}$ as desired. $\blacksquare$
%Problem 5
\newpage
\section{}

%Problem 6
\newpage
\section{}

%Problem 7
\newpage
\section{}

>>>>

%Problem 3
\newpage
\section{}
\paragraph{}
We throw $n^{0.99}$ balls into $n$ bins uniformly at random. Using union bound, the probability that a bin has at least $k$ balls, for $k$ chosen later, is at most 
$${n \choose k}(\frac{1}{n})^k.$$
Using the bound ${n^{0.99}\choose k} <(\frac{n^{0.99}e}{k})^k$ this probability is at most
$$(\frac{n^{0.99}e}{k})^k(\frac{1}{n})^k= \frac{e^k}{n^{0.0k}k^k}\leq \frac{e^k}{k^k}$$
where the last inequality follows since we will choose $k \geq 0$ and since $n\geq 1$ this implies $n^\frac{k}{100} \geq 1$.
Again invoking union bound
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq n\frac{e^k}{k^k} = n(\frac{e}{k})^k.$$ 
Since we want
$$Pr(\exists\text{a bin with at least $k$ balls}) \leq \frac{1}{2}$$
we need to find $k$ such that
$$n(\frac{e}{k})^k \leq \frac{1}{2}.$$

%Problem 4
\newpage
\section{}
\paragraph{}
We always use addition, $+$, over the appropriate field (ie. $\F_n$ for inputs and $\F_m$ for outputs). Our algorithm is very simple. Given input $z$ choose $x \in \{0,\dots, n-1\}$ uniformly at random. Let $y = z-x$. Lookup $F(x)$ and $F(y)$ and return $F(x) + F(y)$.
\paragraph{}
If both $F(x)$ and $F(y)$ are not modified by our adversary then our return value satisfies
$$F(x) + F(y) = F(x+y) = F(z)$$
using the linearity of $F$ and the fact that $x+y = z$. So a necessary condition for our algorithm to fail is that one of $F(x)$ or $F(y)$ was modified by our adversary. For a uniform random element of the table, the probability of modification is at most $\frac{1}{5}$. Hence
$$Pr(F(x)\text{ or } F(y)\text{ was modified}) \leq Pr(F(x) \text{ was modified}) + Pr(F(y)\text{ was modified}) \leq \frac{2}{5}.$$
Therefore the probability of failure is at most
$$\frac{2}{5} < \frac{1}{2}.$$
Hence our algorithm succeeds with probability at least $\frac{1}{2}$ as desired. $\blacksquare$

%Problem 5
\newpage
\section{}

%Problem 6
\newpage
\section{}
\paragraph{}
Let $\cA$ be an algorithm which obtain an $(\epsilon, \frac{1}{4})$-approximation to some parameter $\mu$. That is, when $X$ is the output value of $\cA$,
$$Pr(|X - \mu| \leq \epsilon\mu) \leq \frac{1}{4}.$$
We want to show how to obtain an $(\epsilon,\delta)$-approximation to $\mu$. Suppose we perform $k = \ceil{\log(\frac{1}{\delta})}$ independent runs of $\cA$ and output the median. We may assume $k$ is even, if not replace $k$ with $k+1$. Let $X_1, \dots, X_k$ be random variables each denoting the output value of one run of $\cA$. Instead of ordering them by the order the trials happen, order them sorted by value, i.e.
$$X_1 \leq X_2 \leq \dots\leq X_k.$$
Our $(\epsilon,\delta)$-approximation algorithm outputs $X_\frac{k}{2}$, the median value of our independent trials. Let $X$ denote the output of our new algorithm. We observe that $|X - \mu|\leq \epsilon \mu$ if and only if 
$$|X_i - \mu | \leq \epsilon \mu, \quad \forall i=1 \dots \frac{k}{2}.$$
Therefore
$$Pr(|X-\mu| \leq \epsilon \mu) = \Pi_{i=1}^\frac{k}{2} Pr(|X_i - \mu| \leq \epsilon \mu) \leq \frac{1}{4^\frac{k}{2}}.$$
Since $k = \ceil{\log(\frac{1}{\delta})}$,
$$\frac{1}{4^\frac{k}{2}} \leq \frac{1}{4^{\frac{1}{2}\log(\frac{1}{\delta})}} = \delta.$$
Therefore
$$Pr(|X-\mu| \leq \epsilon \mu) \leq \delta. \blacksquare$$
%Problem 7
\newpage
\section{}
